---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data: Bechdel Test Data

This data is from fivethirtyeight.come, and was originally collected from IMBD movie data in 2017.

The observations in this datasets are various movies, and the variables include: year (year movie was released), imdb (imdb code for movie), title (title of movie), test (whether it passes each criterion of the Bechdel test- contains two named female characters, do these two women speak to each other, do they speak to each other about something other than a man), binary (whether the movie passes the Bechdel test) budget (movie budget), domgross (the domestic gross product of the movie, intgross (the international gross product of the movie), and the budget/GDP of the movies converted to 2013 dollars. 

## Research question: Bechdel Test Data

Do movies that pass the Bechdel test have higher or lower budgets than movies that fail the Bechdel test? Do the movies that pass this tend to gross larger profits than those that fail, and is there a difference between this on the domestic and international scale?

This question is important as it relates to the level of female representation in the film industry, and the value that the film industry places on female actors, characters, and storylines in film. 

Our group is researching the relationship between film budget/profit compared to whether these films being analyzed pass the Bechdel test. We hypothesize that movies that fail the Bechdel test will tend to have lower budgets than movies that pass the Bechdel test. 

There are both quantitative variables and categorical variables at use in our research question. For example, whether a movie passes none, some, or all of the Bechdel test is a categorical variable, while the recorded budget of a movie is a quantitative variable. 

## Literature: Bechdel Test Data:

Our research question sparked from a fivethirtyeight.com article that described how ⅓ of the top 50 movies from 2016 failed the Bechdel Test. The Bechdel Test is a simple test that assesses movies based on three criteria: does a movie have at least two named female characters? Do these named characters have at least one conversation? Is this conversation about something that is not about a man? We were surprised by the number of movies that did not pass this seemingly basic test, even in 2016. This article further explored diversity in the film industry, testing movies on whether the on-set crew were 50% women (of which none of these top 50 movies passed), and explored the racial diversity in the film industry. 

To further explore female representation and characterization in the film industry, we want to investigate the relationship between film budget and profit in regard to whether films pass the Bechdel test. 

(article: https://projects.fivethirtyeight.com/next-bechdel/)\
\
Glimpse of data

```{r}
#| label: load-data-1

bechdel<- read.csv(file = 'data/bechdel_clean.csv')
glimpse(bechdel)
```

# Data 2

## Introduction and data: **Wealth and Income**

-   The source of this dataset is the Urban Institute and the US Census.

-   Overall the U.S. Census Bureau states that there are three methods of data collection-- mail, Internet collection, and collection from state sources.

-   In the dataset, there are 2,916 observations with variables surrounding race, income statistics, and year.

-   For ethics, since each observation is anonymous, there should be no privacy concerns. However, U.S. Census data has historically underrepresented marginalized groups so the analysis may not be a complete reflection of the U.S as a whole.

## Research question

-   Research Question: How has the relationship between race and income distribution changed overtime in the U.S?

-   Overall, this question is important as the U.S has changed dramatically across it\'s history. More specifically, many marginalized groups have been impacted by certain events and policies-- Jim Crow laws, red-lining, the Great Recession, etc. Viewing the long-term and short-term impacts of these events could help improve future policies and actions by the government.

-   This research topic of Race and Wealth is constantly evolving-- incorporating new variables and details as society learns more about the history about the U.S. My hypothesis is that overall marginalized groups wealth have increased throughout history with the distribution within these groups widening.

-   The types of variables in my research question include year (quantitive), race (categorical), income mean (quantitative), income bracket (categorical), and income distribution (quantitative).

    ## Literature

```{=html}
<!-- -->
```
-   <https://www.federalreserve.gov/econres/notes/feds-notes/disparities-in-wealth-by-race-and-ethnicity-in-the-2019-survey-of-consumer-finances-20200928.html> 

-   Overall this article looks at recent trends in wealth between different races in the U.S. It found that Black and Hispanic families\' incomes rose dramatically after experiencing the aftermath Great Recession more severely. Additionally, it found that within every race, wealth was significantly higher for middle-aged families compared to older and younger families. However, the disparity between these subgroups is still apparent. Finally the article details the disparity in retirement account ownership between different races.

-   Overall this research paper looks at many factors contributing to wealth in a short period of time (2007-2019). My analysis however will do the opposite- look at just one factor, race, and see how that impacts income distribution over a long period of time (1963 - 2019).

## Glimpse of data

```{r}
#| label: load-data-wealth-income

income_distribution <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_distribution.csv')

glimpse(income_distribution)
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

# add code here
```
