---
title: "The Relationship between Race and Wealth in U.S. history over Time"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

## Introduction

Research Question: Does a movie's passage of the Bechdel test influence its success?

## Methodology

```{r}
#| label: load-packages
#| message: false
library(tidyverse)
```

```{r}
#| label: load-data-wealth-and-income

bechdel<- read.csv(file = 'data/bechdel_test.csv')
bechdel
```

```{r}
bechdel|>
  ggplot(aes(x = year, y = budget, color = binary )) +
  geom_point()
  ggplot(aes(x = year, y =budget))+
  geom_point() +
  facet_wrap(~binary) +
  geom_smooth(method = lm, se = FALSE)
```

```{r}
bechdel|>
  ggplot(aes(x = year, fill = binary)) +
  geom_bar(position = "fill")
bechdel|>
  filter(year == 1970)
bechdel|>
  ggplot(aes(x = year)) +
  geom_histogram()

```

Visualizations Ideas:

are lower budget or higher budget movies more likely to pass?

```{r}
bechdel|>
summarize(median = median(budget))
low_budget <- bechdel |>
  filter(budget<= 2.8e+07)
high_budget <- bechdel |>
  filter(budget>2.8e+07 )
low_budget|>
  ggplot(aes(x = budget, color = binary)) +
  geom_histogram()
high_budget |>
  ggplot(aes(x = binary)) +
  geom_bar() 


bechdel$budg<- as.factor(ifelse(bechdel$budget<2.8e+07, "low", "high"))
bechdel|>
  ggplot(aes(x = budg, fill = binary)) +
  geom_bar(position = "fill")
bechdel|>
  ggplot(aes(x = budget)) +
  geom_boxplot()
 
```

Income distribution for every race. Facet wrap? or By group.

Stacked bar chart, below poverty and above poverty.

Graph where we manually draw poverty line?

Summary Statistics ideas:
